---
title: "MAS8403"
author: "Harvey Yuan 0077439"
date: '2022-10-31'
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(tidyverse)
library(knitr)
library(magick)

data("penguins")
penguins = na.omit(penguins)
my.student.number = 0077439
set.seed(my.student.number)
my.penguins = penguins[sample(nrow(penguins), 100), ]
mean.penguin <- aggregate(body_mass_g~species+island+sex,my.penguins,mean)
attach(mean.penguin)
sorted.mean <- mean.penguin[order(species),]
detach(mean.penguin)

species.penguin <- aggregate(body_mass_g~species+sex,mean.penguin,mean)

attach(species.penguin)
sorted.species <- species.penguin[order(species),]
detach(species.penguin)

island.mean <- aggregate(body_mass_g~island+sex,my.penguins,mean)

```

## Aim and Objective

The aim of this analysis is to explore and see if there is any
correlation between the which sex the penguins are from and their body
mass index. This will hopefully allow researchers to understand what
causes the difference in weight for the penguins such as environmental
and genetic factors.

From this data, we hopefully will be able to use body mass to
distinguish between male and female penguins without using invasive
procedures. By using analysis methods which are non-invasive for sexing,
less stress will be inflicted onto the penguins which will reduce the
weight fluctuations of the penguins due to stress.

## Technologies and Libraries Used

***R version 4.1.2***

R is a language used for statistical computing and graphics.

***RStudio 2021.09.01 Build 372***

RStudio is an IDE that is used for R. RStudio includes an in-built
console and terminal that allows for direct code executing and also a
pane which allows for files, plots and packages management. Different
libraries can be installed through this pane.

***dplyr version 1.0.7***

The dplyr library allows for the user to use a set of verbs to write
code for common data manipulation steps. It allows the user to use
familiar words when scripting analysis and pre-processing scripts (such
as filter), both for easy understanding of the code and proof reading
(Hadley Wickham, n.d.).

***readr version 2.0.1***

The readr library provides for an easy way to read the csv files that
are being used as part of this analysis. As well as being easier to for
users to understand and read the data, it also makes analysis more
reproducible as base R functions inherit behaviours from the operating
system and environment variables, as such, importing code from one
environment to another using readr will work without issue (Grolemund,
2016).

***ggplot2 version 3.3.5***

ggplot is a system for creating graphics (where the data is a selected
data frame). The data is either provided by the user or created in
script, which allows for said user to iteratively add new layers,
components and functionality. ggplot will be used to demonstrate the
analysis as part of this investigation.

\newpage

## Data Preparation and Analysis

As part of the project, the prepared dataset which provided the data on
the penguins was taken from the "palmerpenguins" package in R. The
dataset contains data on 333 penguins with a large number of different
variables but for the investigation, we're going to wrangle and analyse
the following data:

• species - species of the penguin (Adelie, Chinstrap or Gentoo)

• island - island the penguin lives (Biscoe, Dream or Torgerson)

• body mass g - penguin's body mass (in grams)

• sex - sex of the penguin (male or female)

For the investigation, a sample of 100 penguins were used and the
collected data of these penguins were analysed.

With the understanding of this data and the aim of this investigation,
in the first instance, it was decided that the first thing to look at is
the sample mean of the body masses of the penguins aggregated by
species, island and sex. The aggregation of the data provides the sample
mean which is a good estimator of the population mean.

```{r echo=FALSE, fig.align = 'center'}

options(repr.plot.width = 1, repr.plot.height =1)

g <- ggplot(my.penguins, aes(x = body_mass_g)) +
    stat_function(
        fun = dnorm,
        args = with(my.penguins, c(mean = mean(body_mass_g), sd = sd(body_mass_g)))
    ) + xlim(2000, 6500)+theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
        )

ggsave(g, file = "normdis.pdf", width = 4, height = 4)

include_graphics(paste0(getwd(),"normdis.pdf")) 

```

By plotting the body mass, we can tell that the data is normally
distributed. This means that the mass of the penguins distribute
symmetrically around the mean mass of the penguins and that a majority
of the penguins fall within the ±1.96 standard deviations of the sample
mean.

\newpage
```{r echo=FALSE}
kable(sorted.mean, caption="Aggregated Mean Body Mass")


```

The above table shows the sample mean of the 100 penguins sorted by
species.

The mean body mass of the female penguins is 4,219g whereas the mean
body mass of the male penguins come to 3,818g. Comparing the body mass
of the penguins by species reveals that the female Adelie penguins, on
average, have 168g less body mass whereas for the other two species, the
females have a higher body mass of 224g and 657g respectively for the
Chinstrap and Gentoo species.

It is worth noting that for both the Chinstrap and Gentoo species of
penguins, they only appear on one island each. The Chinstrap species
only appear on the Dream island whereas the Gentoo species only appear
on the Biscoe island. This is true for the population mean and by
extension, also true for the sample that we have.

```{r echo=FALSE}

kable(species.penguin, caption="Mean Body Mass by Species")
```

In table 2, the data shows that for the Adelie species of penguin, the
male penguins have a larger body mass regardless of which island they
are from by approximately 685g.As for the Chinstrap and Gentoo penguins,
they data also shows that the male penguins have a larger mean body mass
than their female equivalents by approximately 224g for the Chinstrap
and 657g for the Gentoo.

```{r echo=FALSE}

kable(island.mean, caption="Mean Body Mass by Island")

```

This realisation is confirmed by table 3, as shown above. Table 3 shows
the mean sample mass for the penguins aggregated by the islands they are
from. This suggests that, on a whole, that the male population of
penguins have a higher body mass than the female penguins.

By using the values from tables 1,2 and 3, we can calculate the
confidence interval, by the means of the t distribution, of the
population mean with 95% confidence that it would fall within the ranges
of:


3842.278477 \< $\mu$ (female) \< 3868.435809

4510.617446 \< $\mu$ (male) \< 4541.106691

By calculating the 95% confidence interval of our estimator, the sample
mean, we can say that the population mean for body mass will fall within
the given ranges for both sexes 95% of the time.

From the data and analysis, it seems as though using the body mass of
the penguins is a good way to distinguish the sexes of the penguins.
This is assuming that the penguins are not abnormally large or
abnormally small for the sex of the penguin. This means that for any one
individual penguin, if their body mass lies within ±1.96 standard
deviations of the sample mean range as shown above, they are likely to
fall within they sex of penguin.

To analyse if the island which the penguin is on has any significant
impact on it's physical characteristic, we would need to discount the
Chinstrap and Gentoo species because the Adelie species is the only
observed species which are present across all three islands.

As part of the analysis to identify if the islands made any significant
impact on any physical characteristic, the confidence internals were
worked out of the sample body mass for the male and female penguins of
the Adelie penguins. By calculating this, we would be 95% confident that
the population body mass for the male and female penguins lie within the
range. The ranges of the upper and lower mean of the Adelie penguins are
as follows:

3305.007179 \< $\mu$ (female) \< 3532.492821

3949.159351 \< $\mu$ (male) \< 4252.923982


Comparing the mean body mass of the Adelie species across the islands, it
doesn't look like the island affects the body mass as the means of the
body mass across the species varies from the range of the confidence
interval.

```{r echo=FALSE}
filtered <- filter(my.penguins, species =="Adelie")

mean.filtered <- aggregate(body_mass_g~species+island+sex,filtered,mean)

kable(mean.filtered, caption="Adelie by Sex and Island")
```

From table 4, all body mass sample means for the Adelie penguins fall within the population mean calculated using the confidence intervals. This suggests that which island the Adelie penguins are on doesn't affect the body mass.

\newpage

## Critical Evaluation

The aim of this investigation was to determine if there is any correlation between which sex the penguins are and their body mass index and hopefully be able to determine the sex of the penguins using non-invasive procedure. 

It was chosen that the body mass would be used as part of the physical trait to evaluate as generally in nature, male versions of the same species are larger than their female counterparts. The data was plotted to show that the body mass of the sample was normally distributed, which allowed the for the population mean to be approximated to fall within the ranges 95% of the time.

The body mass of the penguins where aggregated by species, island and sex in the first instance which showed that only the Adelie penguin were present on all three islands. The body mass of the penguins were also evaluated against their species and sex. This showed that the across all three species, the male penguins were on average approximately 520g larger than the female penguins.

Lastly, as the Adelie penguins were the only penguins to span across all three islands, they were the only penguins used in the investigation for if which island penguins come from have any significant impact on any characteristic. This meant that the Chinstrap and Gentoo penguins did not affect the sample mean of the body mass and skew it to negatively impact the Dream and Biscoe islands.

To improve this investigation in the future, the body mass of an individual penguin should be evaluated against the calculated figures above and the probability of the sex of the penguin could be determined.

Another way to improve this investigation would be to evaluate the body mass in conjunction with another physical characteristic such as bill depth to allow for more accurate sexing of the penguins. By doing so, the probablity fitting both criteria of male and female characteristics are smaller if the more traits that are compared.

\newpage

## *Bibliography*

Hadley Wickham, R. F. (n.d.). dplyr. Retrieved from dplyr part of the
tidyverse 1.0.7: <https://dplyr>. tidyverse.org/ (Last accessed 1st of
November 2022)

Grolemund, H. W. (2016, December). Data Import. Retrieved from R for
Data Science: <https://r4ds.had>. co.nz/data-import.html (Last accessed
1st of November 2022)
